The Language of Generalization

### 2018-10-26 paper/comgen-paper.tex
tweak wording in causal intro
update readme; remove packrat; update tex

\documentclass[english,floatsintext,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={The language of generalization: Probability, vagueness, and interaction},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{The language of generalization: Probability, vagueness, and interaction}

  \shorttitle{The language of generalization}

  \author{Michael Henry Tessler\textsuperscript{1}~\& Noah D. Goodman\textsuperscript{1}}

  \def\affdep{{"", ""}}%
  \def\affcity{{"", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Department of Psychology, Stanford University  }

  \authornote{
    \newcounter{author}
    This manuscript is currently in prep. Comments or suggestions should be
    directed to MHT.

                      Correspondence concerning this article should be addressed to Michael Henry Tessler, 450 Serra Mall, Bldg. 420, Rm. 316, Stanford, CA 94305. E-mail: \href{mailto:mtessler@stanford.edu}{\nolinkurl{mtessler@stanford.edu}}
                          }

  \abstract{Generalizations are central to human understanding and language provides
simple ways to convey them (e.g., ``Birds fly''). The language of
generalization is ubiquitous in everyday discourse and child-directed
speech. Yet the meaning of these expressions is logically puzzling
(e.g., not all birds fly) and has resisted precise formalization. The
major issue in formalizing generalization in language is determining
which statements are true or which are false. Using an
information-theoretic probabilistic model, we explore the hypothesis
that the meaning of these linguistic expressions is \emph{simple but
underspecified}, and that general communicative principles can be used
to establish a more precise meaning in context. To test this theory, we
examine endorsements of generalizations about three different domains:
categories (\emph{generic language} e.g., ``Birds fly''), events
(\emph{habitual language} e.g., ``John runs''), and causes (\emph{causal
language} e.g., ``Staring at the sun makes you go blind''). Across these
diverse domains, we find that our model explains the wide variance in
human endorsements of language conveying generalization, while simpler
models fall short. Ours is the first formal theory that makes precise,
quantitative predictions about human understanding of the language of
generalization. The results demonstrate that the context-sensitivity of
such language emerges from the interaction of an underspecified meaning
with diverse beliefs about properties.}
  \keywords{genericity, generalization, generics, pragmatics, semantics, bayesian
modeling \\

    \indent Word count: 20000
  }

  \usepackage{tabularx}
  \usepackage{multicol}
  \usepackage{wrapfig}
  \usepackage{caption}
  \usepackage{booktabs}
  \usepackage{amsmath}
  \usepackage{graphicx}
  \usepackage{subcaption}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{multirow}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

\newcommand{\denote}[1]{\mbox{ $[\![ #1 ]\!]$}}
\newcommand*\diff{\mathop{}\!\mathrm{d}}

\definecolor{Red}{RGB}{255,0,0} \definecolor{Green}{RGB}{10,200,100}
\definecolor{Blue}{RGB}{10,100,200}

@@

\section{Introduction}\label{introduction}

Learning that an object tends to have a property, an entity tends to
exhibit a behavior, or a cause tends to produce an effect can be crucial
to thrive in our open-ended world. Yet such knowledge can be difficult
to acquire: Relevant observations may be costly (e.g., learning that a
plant is poisonous) or rare (e.g., understanding that lightning strikes
tall objects). It thus is important that language allow us to
communicate such \emph{generalizations} to each other. By sharing
generalizations, we flourish collectively without individually needing
to taste potentially-poisonous plants or personally witness many
lightning strikes. Flexibly communicating generalizations from one
generation to the next may be a foundational mechanism behind the
faithful transmission of knowledge necessary for cumulative cultural
evolution (Henrich, 2015; Tomasello, 1999).

The language of generalization is often studied via \emph{generic
language} (or, \emph{generics}): simple statements predicating
properties of categories (e.g., ``Dogs are friendly''; Carlson, 1977;
Cohen, 1999; Leslie, 2007; Nickel, 2008). Generics are ubiquitous in
everyday conversation as well as in child-directed speech (Gelman,
Goetz, Sarnecka, \& Flukes, 2008). In contrast to \emph{particular
statements} (e.g., \enquote{Rufus is friendly}) which describe
individual entities or situations and can be verified through
observation, generic statements refer to inherently unobservable
categories (e.g., the category of \textsc{dogs}) and convey information
that extends beyond the present context, a fact which children as young
as 2 appreciate (Cimpian \& Markman, 2008). Generics are just one case
in the language of generalization, however; events can be described in
generalization (\emph{habitual language}; e.g., \enquote{Mary swims
after work.}) as well as causal relationships (\emph{causal language};
e.g., \enquote{Staring at the sun makes you go blind.}). Every language,
it is believed, can express generalization (Behrens, 2005; Carlson \&
Pelletier, 1995) and is central to the growth of conceptual knowledge
(Gelman, 2004). Considering generalization as a singular phenomenon in
language, thus, promises to provide understanding of a wide variety of
linguistic expressions with powerful implications.

Despite their ubiquity and relative simplicity, there have been few
formal models of the language of generalization and none that have made
precise, quantitative predictions to match human intuitions. The major
challenge in formalization is determining \emph{what makes the
statements true or false}. When we hear \enquote{Dogs bark}, it sounds
like a universal statement, as in \enquote{All dogs bark}. But the
property does not apply to all dogs: Basenji dogs are barkless\footnote{\url{http://strangesounds.org/2014/09/basenji-dogs-basenjis-dont-bark-howl.html}},
for example. The generalization could signal something weaker like
\enquote{Most dogs bark}. However, most mosquitos do not carry malaria
(in fact, only a tiny fraction do), but \enquote{Mosquitos carry
malaria} is intuitively true. The only quantifier that would permit such
a statement would be \enquote{some} (e.g., \enquote{Some mosquitos carry
malaria}). But weakening the generalization to mean \enquote{some} is
too permissive: some robins are female (in fact, half of them are), but
asserting \enquote{Robins are female} seems odd. Even more perplexing:
Those very same female robins lay eggs, and \enquote{Robins lay eggs} is
a perfectly fine thing to say.

These observations have led to the conclusion that the prevalence of the
feature (or the probability of a member of the category having the
property e.g., \(P(x \text{ barks} \mid x \text{ is a dog})\)) is not
the only parameter in the meaning of generalizations. Conceptual
accounts of generics emphasize the structure of \emph{generic knowledge}
(Prasada, 2000), and view generic utterances as the way of expressing
special, mental relationships between kinds and properties (Leslie,
2008; Prasada, Hennefield, \& Otap, 2012). \enquote{Bishops move
diagonally} not because bishops tend to move diagonally but rather
because those are the rules of the game. It's been suggested that the
mind has cognitive primitives that directly dictate what generalizations
are true or false. One such primitive is the ability to priviledge
\emph{characteristic features}: For an animal, their mode of
reproduction is a characteristic feature. Thus, whatever their mode of
reproduction may be will make a true generic statement (e.g.,
\enquote{Robins lay eggs}), regardless of the actual prevalence of the
feature. \emph{Horrifying} or \emph{striking} information is also
priviledged and creates true generics, which is why \enquote{Mosquitos
carry malaria} is true

Yet, generics convey more than just rich, conceptual relations (Nickel,
2008, 2016; Sterken, 2015). Purely arbitrary relationships can result in
true generics. \enquote{Barns are red} because barn-owners tend to paint
their barns red; were barn-owners to decide tomorrow to switch to blue,
then \enquote{Barns are red} would be false. Ravens and toasters have no
priviledged relationship to each other, but the statement
\enquote{Ravens are bigger than toasters} is still a true
generalization. Conceptual accounts must be supplemented with the
proviso that, in some situations, the actual prevalence of the feature
matters for making generalizations true.

What we are left with is hybrid theories, which try to classify
different kinds of generalizations and the various inferences they
support (e.g., Prasada, Khemlani, Leslie, \& Glucksberg, 2013). But
insofar as there is a single class of linguistic expressions that convey
generalization, there should be something common to them all: a general,
unified semantic core. In this paper, we propose such a semantic core
for the language of generalization. To do so, we look to probability,
the universal currency of belief and a useful representation for human
generalization from observations (Shepard, 1987; Tenenbaum \& Griffiths,
2001). Given that certain probabilities are the result of inductive
generalization, it would make sense that these are the object of
communication in the language of generalization. To formalize a model of
endorsement (i.e., what makes the statements true or false), we consider
the information-theoretic function of a generalization by explicitly
modeling the background knowledge a language interpreter would use when
understanding such language. We create an endorsement model based on
this interpretation hypothesis, providing the first formal theory that
makes quantitative predictions about what makes generalizations in
language true or false. The assumptions of the theory, and in turn, the
conditions under which the theory may be meaningfully tested are made
explicit by its instantiation in a formal model. Though concepts such as
\emph{striking properties} are not explicitly formalized in our model,
our general cognitive modeling framework provides natural avenues for
future integration.

The paper is organized as follows. In Section 1, we formalize the
hypothesis that the core of meaning of a generalization in language is
simple but \emph{underspecified}, articulate a thory of relevant
background knowledge used to understand a generalization in language,
and develop a computational model of endorsement around these
components. The endorsement model's predictions depend upon the
background knowledge and the probability communicated, which we
demonstrate via simulation. Sections 2 - 4 present our three empirical
case studies: generalizations about categories (\emph{generic
language}), events (\emph{habitual language}), and causes (\emph{causal
language}). Across these three case studies, we both measure and
manipulate background knowledge and the probability communicated,
measuring their effects on endorsement. We find a very strong agreement
of our model's predictions to the full spectrum of human elicited
endorsements, where alterantive models fall short. This work opens the
door to understanding generalizations in pragmatic language
understanding contexts and building models of learning from language.

\section{Computational Framework}\label{computational-framework}

Generalizations are used to make predictions about properties of
instances that an agent has yet to experience (Hume, 1888). People
readily predict that a new dog will have fur, drinking a new cup of
coffee will cause jitters, and a new day will find certain people
typing. In each case, we assign a specific exemplar \(x\) to a category
\(k\), and make a prediction that it will have property \(f\). This
prediction can be described by a conditional probability:
\(p_{fk} = P(x \in f \mid x \in k)\), the probability that \(x\) will
have \(f\) (formally, be in the set of things that have \(f\)) given
that it is in \(k\), which we will refer to as the \textbf{prevalence}.
The targets of our predictions can vary widely: They may be objects (a
dog), events (a coffee drink), or more fine grained types (a person on a
given day). The properties also vary (e.g., having fur, causing jitters,
typing). Yet the mathematical description of the inductive inference is
always given by the probability \(p_{fk}\).

How does one acquire such an inductive belief? Observing instances of
the category \(k\) and noting how many of them have \(f\) can support
prediction to new instances when integrated with one's prior
expectations about the property (cf., Gelman, 2003; Keil, 1992). For
example, if you observe creatures that \textsc{have two legs}, you might
infer that all creatures of that category have two legs. Observing new
creatures with \textsc{broken wings}, however, leads to a much weaker
prediction about other instances of the category having the property
(Nisbett, Krantz, Jepson, \& Kunda, 1983). This kind of domain-specific
knowledge about properties can be represented formally as a probability
distribution over the prevalence \(p_{f}\): \(P(p_{f})\) (Kemp \&
Tenenbaum, 2008), and observational data \(d\) can update one's beliefs
via Bayes' Rule:
\(P(p_{f} \mid d) \propto P(d \mid p_{f}) \cdot P(p_{f})\)

Learning from observational evidence can only take you so far, though.
To learn about properties that are costly to observe (e.g.,
\emph{staring at the sun makes you go blind}) or events that are
statistically rare (e.g., \emph{lightning strikes tall objects}), one
would much rather learn from others. Fortunately, language has simple
ways of encoding generalizations.

\subsection{Generalization from
language}\label{generalization-from-language}

The language of generalization manifests when a property is predicated
of a category: \enquote{Ks F} (e.g., \enquote{Dogs bark},
\enquote{Staring at the sun makes you go blind}). Generalization from
observations can be described as a predictive probability; it is
intuitive then to posit that same quantity will be at the core of a
semantic theory of the language of generalization. To connect to the
semantic tools of \emph{truth values} (Montague, 1973), we must relate
the scalar quantity of probability to a Boolean value. One simple way is
with a \emph{threshold semantics}: An utterance is true if the relevant
quantity is above a threshold. A threshold semantics for a
generalization about category K with property F would then apply to the
\emph{prevalence} \(p_{fk}\):
\(\denote{K \text{ } F}(p_{fk}, \theta) = \{p_{fk} >\theta\}\)

Various attempts of this kind have been made before (e.g., Cohen, 1999).
We distinguish our proposal from previous ones by emphasizing the
cognitive underpinnings of \(p_{fk}\), as well as the formalization into
a model that makes precise, quantitative predictions about human
endorsements. \(p_{fk}\) describes a latent belief and should not be
confused with \emph{frequency}, which can inform a belief but is not
itself the belief.

\subsubsection{Literal interpretation
model}\label{literal-interpretation-model}

Quantifier statements can be described using a similar threshold
semantics:
\(\denote{some \text{ } K \text{ } F}(p_{fk}) = \{p_{fk} > 0\}; \denote{all \text{ } K \text{ } F} = \{p_{fk} = 1\}\).
So what threshold \(\theta\)? should be used for the generalizations?
Unlike quantified language which have fixed-thresholds, we posit that
the language of generalizations is \emph{underspecfied} with respect to
the actual truth-functional threshold \(\theta\). That is, we
hypothesize that \(\theta\) is in part determined by contextual, in a
way analagous to how gradable adjectives like \enquote{tall} have
contextually-determined thresholds (e.g., what counts as ``tall'' for a
four-year-old is different than what counts as ``tall'' for an adult;
Kennedy, 2007; Lassiter \& Goodman, 2013).

>>>>

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The Language of Generalization},
            pdfauthor={Michael Henry Tessler~\& Noah D. Goodman},
            pdfkeywords={genericity, generalization, generics, pragmatics, semantics, Bayesian
modeling},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{The Language of Generalization}
    \author{Michael Henry Tessler\textsuperscript{1}~\& Noah D.
Goodman\textsuperscript{1}}
    \date{}

\shorttitle{The language of generalization}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\keywords{genericity, generalization, generics, pragmatics, semantics, Bayesian modeling\newline\indent Word count: 27578}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
\usepackage{tabularx}
\usepackage{multicol}
\usepackage{wrapfig}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{xcolor}

\authornote{This paper is currently under review. Comments or
suggestions are appreciated and should be directed to MHT.

Correspondence concerning this article should be addressed to Michael
Henry Tessler, 450 Serra Mall, Bldg. 420, Rm. 316, Stanford, CA 94305.
E-mail:
\href{mailto:tessler@mit.edu}{\nolinkurl{tessler@mit.edu}}}

\abstract{
Language provides simple ways of communicating generalizable knowledge
to each other (e.g., ``Birds fly'', ``John hikes'', ``Fire makes
smoke''). Though found in every language and emerging early in
development, the language of generalization is philosphically puzzling
and has resisted precise formalization. Here, we propose the first
formal account of generalizations conveyed with language that makes
quantitative predictions about human understanding. We test our model in
three diverse domains: generalizations about categories (generic
language), events (habitual language), and causes (causal language). The
model explains the gradience in human endorsement through the interplay
between a simple truth-conditional semantic theory and diverse beliefs
about properties, formalized in a probabilistic model of language
understanding. This work opens the door to understanding precisely how
abstract knowledge is learned from language.

}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

\newcommand{\denote}[1]{\mbox{ $[\![ #1 ]\!]$}}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\definecolor{Red}{RGB}{255,0,0}
\definecolor{Green}{RGB}{10,200,100}
\definecolor{Blue}{RGB}{10,100,200}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Knowledge that extends beyond the present context is crucial to thrive
in our open-ended, dynamic world. Yet such knowledge can be difficult to
extract from the environment: The relevant observations may be costly
(e.g., learning that a plant is poisonous) or rare (e.g., understanding
that lightning strikes tall objects). Fortunately, we are not limited to
acquiring generalizations on our own; language allow us to communicate
generalizations to each other. By sharing generalizable knowledge, we
flourish collectively without individually needing to taste
potentially-poisonous plants or personally witness lightning strikes.
Being able to flexibly communicate generalizations from one generation
to the next supports the faithful transmission of knowledge necessary
for culture to cumulatively evolve (Henrich, 2015; Tomasello, 1999).

The \emph{language of generalization} covers a diverse swath of natural
language expressions. Generic language conveys generalizations about
categories (e.g., ``Dogs have four legs''; Carlson, 1977; Cohen, 1999;
Leslie, 2007; Nickel, 2008) and is the most well-studied case
generalizations in language.\footnote{Some writers refer to the all of
  the \emph{language of generalization} as \enquote{generic language} or
  generics. In both the empirical and theoretical literatures, however,
  analysis and experiments often focus only on generalizations about
  categories. We use the more narrow-scoped terms (e.g., generics,
  habituals, and causals) to highlight the diversity of semantic types
  being predicated.} In contrast to statements about concrete
individuals (e.g., \enquote{Rufus has four legs}), generic statements
refer to inherently unobservable categories (e.g., the category of
\textsc{dog}) and convey information that extends beyond the present
context, a fact which children as young as 2 appreciate (Cimpian \&
Markman, 2008). Simple events (e.g., \enquote{John ran yesterday}) can
be generalized into habitual sentences (e.g., \enquote{John runs}), and
even events of complex inferential types such as actual causal events
(e.g., \enquote{The fire caused the smoke}) can be described in
generalization (e.g., \enquote{Fire causes smoke}).

Understanding the language of generalization is a project with
far-reaching implications. The language of generalization is ubiquitous
in everyday conversation, is found in every language (Behrens, 2005;
Carlson \& Pelletier, 1995), and conveys rich meanings, impacting
motivation (Cimpian, Arce, Dweck, \& Markman, 2007), transmitting
stereotyped beliefs about social categories (Rhodes, Leslie, \& Tworek,
2012), and making meaning from experience (Orvell, Kross, \& Gelman,
2017). It is highly-prevalent in child-directed speech (Gelman, Goetz,
Sarnecka, \& Flukes, 2008) and its ability to refer to abstractions
beyond the present context suggests its centrality to the growth of
conceptual knowledge (Gelman, 2004).

Despite its ability to convey abstract knowledge, its ubiquity in
discourse, and its relative morphosyntactic simplicity, the language of
generalization displays subtle context-sensitivities that make it
difficult to formalize. \enquote{Robins lay eggs} sounds true and
\enquote{Robins are female} does not. Yet in each case, only 50\% of the
category has the property (i.e., only the females lay eggs).
\enquote{Mosquitos carry malaria} sounds true despite malaria being
present in only a tiny fraction of mosquitos. Even more perplexing:
\enquote{Supreme Court Justices have even social security numbers} is
thought to be intuitively a false generalization even if it were the
case that on the current bench, nine out of nine justices had even
social securtity numbers (i.e., when the sentence ``All Supreme Court
Justices have even social security numbers'' is true; Cohen, 1999).
Similar context-sensitivity can be observed with habituals:
\enquote{Mary climbs mountains} could imply a few mountain climbs a year
for Mary, but \enquote{John runs} would be infelicitious if John went
for a run three times last year.

These observations have led some to conclude that the literal meaning of
a generic statement (and by analogy, other generalizations in language)
involves more than just the number of members of the category who have
the property, otherwise known as the \emph{prevalence} of the feature in
the category. Instead, theorists have argued that generics are not
treatable by the standard tools of truth-functional semantics (Montague,
1973), but rather should be thought of semantically as a direct,
linguistic manifestation of abstract relations between kinds and
properties (Leslie, 2008; Prasada, 2000; Prasada, Hennefield, \& Otap,
2012). For example, the statements \enquote{Bishops move diagonally} or
\enquote{The Speaker of the House succeeds the Vice President} are true
not because of a tendency on behalf of instances of a category to
actually uphold the property, but rather the existence of a conceptual
relationship (e.g., what it means, in the game of chess, to be a
bishop). True generics can be supported by different underlying types of
category--property relations (e.g., principled vs.~statistical
connections; Prasada \& Dillingham, 2006), and thus support
qualitatively different inferences (e.g., ``Being striped is one aspect
of being a tiger'' is generally endorsed, while ``Carrying malaria is
one aspect of being a mosquito'' is not, even while ``Tigers are
striped'' and ``Mosquitos carry malaria'' are both intuitively true;
Prasada, Khemlani, Leslie, \& Glucksberg, 2013). This conceptual view of
generics has been influential in psychology, because it predicts
qualitative differences between different kinds of generics.

Insofar as there is a single class of linguistic expressions that convey
generalization, however, there should be something common to them all: a
literal meaning that unifies generic, habitual, and causal language. In
this paper, we propose such a semantic core based on prevalence or
probability and formalized using the tools of truth-functional semantics
(Cohen, 1999; Montague, 1973). A semantics based on probability will not
be enough to subtle sensitivities to context that the language of
generalization exhibit. We propose that the meaning of generalizations
is underspecified or \emph{vague}, and that listeners derive a more
precise interpretation in context using world knowledge.

The fact that generics are vague does not preclude them from being
treated with formal models. We draw upon the tools of Bayesian models of
cognition (Tenenbaum, Kemp, Griffiths, \& Goodman, 2011) to formalize
the vagueness and context-dependence of generic language (Frank \&
Goodman, 2012; Goodman \& Lassiter, 2015). The Bayesian model separates
the semantics of an utterance conveying a generalization from the world
knowledge a listener would use to interpret the utterance, a key
theoretical advancement beyond previous accounts. This formal model is
the first of its kind to make quantitative predictions about human
understanding of generalization in language.

The paper is organized as follows. In the next section, we describe our
computational framework for interpreting generalizations and our precise
model for endorsing such statements (i.e., a model of truth judgments).
To illustrate how the model works, we then work through a number of
standard examples from the linguistics literature that any model of
generics should be able to accomodate. The third section discusses the
relationship of our model to previous accounts of generics from the
linguistics and philosophical literatures. These theoretical sections
are followed by three empirical case studies: generalizations about
categories (\emph{generic language}), events (\emph{habitual language}),
and causes (\emph{causal language}). In the case study of generics, we
use measure relevant background knowledge and prevalence to predict
endorsements of familiar generic statements (e.g., \enquote{Robins lay
eggs}). In the study of habituals, we measure background knowledge but
manipulate the prevalence or frequency of the event to predict
endorsements of habitual statements about novel agents (e.g.,
\enquote{John runs} given that he's run a certain number of times in the
past). In this case study, we also further examine the nature of the
relevant probabilities for endorsing generalizations, asking whether it
matters how often John has run in the past (past frequency) or how often
a speaker expects him to run in the future (predictive probability).
Finally, in our last case study, we manipulate background knowledge to
show its causal influence on endorsing generalizations about novel
causal events (e.g., \enquote{Herb X makes wugs sleepy}). We compare our
model to two previously articulated quantitative models of generics as
well as a lesioned-version of our model. In all cases, we find a very
strong agreement of our model's predictions to human elicited
endorsements, where the simpler models fall short. We conclude our paper
by clarifying some of the theoretical claims of the model and discuss
open-questions for this approach.

\hypertarget{computational-framework}{%
\section{Computational Framework}\label{computational-framework}}

Generalizations are used to make predictions about events or properties
of instances that an agent has yet to experience (Hume, 1888). People
readily predict that the next dog they encounter will have four legs,
drinking another cup of coffee will cause jitters, and a new day will
find the people that we know doing what they habitually do. In each
case, we assign a specific exemplar \(x\) to a category \(k\), and make
a prediction that it will have feature \(f\). This prediction can be
described by a conditional probability: \(P(x \in f \mid x \in k)\), the
probability that \(x\) will have \(f\) given that it is in \(k\)
(formally, an instance of \(k\) will be in the set of things that have
\(f\)), which we will refer to as the \emph{prevalence} and for
convenience write as \(p = P(f \mid k)\). Our \emph{prevalence} \(p\) is
a \emph{prevalence in the mind}, a latent belief that a future instance
of a category would have a particular property; others might call this
projectibility (Goodman, 1955). The targets of our predictions can vary
widely: They may be objects (a dog), events (an instance of
coffee-drinking), or more ad-hoc types (a person on a particular day).
The properties also may vary (e.g., having fur, causing jitters, going
to the gym). Yet the mathematical description of the inductive belief is
always given by a probability \(P(f \mid k)\).

Probability is a useful representation for human generalization from
observations (Shepard, 1987). If you observe several \emph{wugs} (a
novel category) that have two legs, you might infer that all wugs have
two legs. But not all properties have such strong projectibility: Seeing
a wug with broken wings tells you comparatively less about other wugs
having broken wings (Nisbett, Krantz, Jepson, \& Kunda, 1983). Abstract,
potentially domain-specific, beliefs about the projectibility of
different properties can be represented by a probability distribution
over the prevalence \(P(p)\) (Kemp, Perfors, \& Tenenbaum, 2007). By
assuming some generative process that could produce one's observations
\(o\) (i.e., a likelihood function \(P(o \mid p)\)), Bayes' Theorem
provides the mathematically correct way to update one's prior beliefs
from observations \(o\): \(P(o \mid d) \propto P(o \mid p) \cdot P(p)\)
(Tenenbaum \& Griffiths, 2001).

Observational data is not always available, however. Instead, we must
listen to others to learn about properties that are costly to observe
(e.g., \emph{staring at the sun makes you go blind}), events that are
statistically unlikely (e.g., \emph{lightning strikes tall objects}), or
any aspect of the world that we have yet to experience. Fortunately,
language provides simple ways of communicating generalizations.

\hypertarget{communicating-generalizations}{%
\subsection{Communicating
generalizations}\label{communicating-generalizations}}
